A pipeline to analyse time-course gene expression data

### 2020-08-11 manuscript.tex
Deploy NelleV/2019timecourse-rnaseq-pipeline to github.com/NelleV/2019timecourse-rnaseq-pipeline.git:gh-pages

The manuscript was last compile on 2020-08-03 07:51:19.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{BiocManager}\OperatorTok{::}\KeywordTok{install}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"NMF"}\NormalTok{,}\StringTok{"ggfortify"}\NormalTok{,}\StringTok{"topGO"}\NormalTok{,}\StringTok{"biomaRt"}\NormalTok{,}\StringTok{"KEGGprofile"}\NormalTok{,}\StringTok{"BiocWorkflowTools"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
- `scores`: the matrix of shape `n_cluster` x `n_genes`,
  containing for each gene and each cluster, the goodness of fit as
  described above.
- `labels`: the labels for all of the genes with a
  sufficiently good goodness-of-fit score.
- `score_cutoff`: the cutoff used on `scores` to determine whether to 
  assign a label
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get the best score and best label for all of the genes}
\CommentTok{# This time without filtering labels}
\CommentTok{# We can give the previous calculated scores to `previous_scores` to save time}
\NormalTok{unfiltered_scores<-}\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(moanin_model,kmeans_clusters,}\DataTypeTok{proportion_genes_to_label=}\DecValTok{1}\NormalTok{, }\DataTypeTok{previous_scores=}\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{scores)}
\NormalTok{best_score =}\StringTok{ }\KeywordTok{rowMin}\NormalTok{(scores_and_labels}\OperatorTok{$}\NormalTok{scores)}
\NormalTok{best_label =}\StringTok{ }\NormalTok{unfiltered_scores}\OperatorTok{$}\NormalTok{labels}

@@

\caption{Heatmap of genes in Cluster 8}\label{fig:heatmapCluster8}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  \textbf{Elbow method}. First introduced in 1953 by Thorndike \citep{thorndike:who},
  the elbow methods looks at the total with cluster sum of squares
  as a function of the number of clusters (WCSS). When adding clusters
  doesn't decrease the WCSS by a sufficient
  amount, one can consider stopping. This method thus provides visual aid to
  the user to choose the number of clusters, but often the ``elbow'' is hard to
  see on real data, where the number of clusters is not clearly defined.
\item
  \textbf{Silhouette method}. Similarly to the Elbow method, the Silhouette method
  refers to a method of validation of consistency within clusters, and
  provides visual aid to choose the number of clusters.
\item
  \textbf{Stability methods} Stability methods are more computationally intensive
  than any other method, as they rely on assessing the stability of the
  clustering for every \(k\) to a small randomization of the data. The user is
  then invited to choose the number of cluster based on a number of similarity
  measures.
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{subsample_proportion =}\StringTok{ }\FloatTok{0.8}
\NormalTok{indices =}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\KeywordTok{dim}\NormalTok{(de_moanin_model)[}\DecValTok{1}\NormalTok{], n_genes }\OperatorTok{*}\StringTok{ }\NormalTok{subsample_proportion, }\DataTypeTok{replace=}\OtherTok{FALSE}\NormalTok{)}

@@

\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & B1 & B2 & B3 & B4 & B5 & B6 & B7 & B8 & B9 & B10 & B11 & B12 & B13 & B14 & B15 & B16 & B17 & B18 & B19 & B20 & B21 & B22 & B23 & B24 & B25\\
\hline
AA204140 & 5 & 5 & 5 & 5 & 5 & 1 & 3 & 2 & 1 & 3 & 1 & 3 & 3 & 1 & 5 & 1 & 1 & 4 & 3 & 5 & 4 & 2 & 4 & 1 & 4\\
\hline
AB008911 & 5 & 5 & 1 & 4 & 1 & 2 & 3 & 3 & 2 & 3 & 3 & 3 & 5 & 1 & 5 & 1 & 3 & 4 & 4 & 5 & 5 & 2 & 4 & 1 & 4\\
\hline
AB010313 & 3 & 3 & 1 & 4 & 1 & 2 & 4 & 3 & 2 & 5 & 3 & 1 & 5 & 3 & 3 & 2 & 3 & 5 & 4 & 2 & 5 & 3 & 3 & 4 & 2\\
\hline
AB010342 & 5 & 3 & 1 & 5 & 1 & 1 & 3 & 3 & 1 & 3 & 3 & 3 & 5 & 1 & 5 & 1 & 1 & 4 & 3 & 5 & 4 & 2 & 3 & 4 & 4\\
\hline
AB011473 & 2 & 1 & 2 & 1 & 3 & 3 & 2 & 5 & 5 & 2 & 4 & 4 & 2 & 5 & 2 & 5 & 5 & 1 & 1 & 1 & 2 & 5 & 1 & 2 & 1\\
\hline
AB099518 & 2 & 1 & 2 & 1 & 3 & 3 & 2 & 5 & 5 & 2 & 4 & 4 & 2 & 5 & 2 & 5 & 5 & 1 & 1 & 1 & 2 & 5 & 1 & 2 & 1\\
\hline
\end{tabular}

@@

\begin{tabular}{llrr}
\toprule
  & Pathway & Percentage & Adj. p-value\\
\midrule
04060 & Cytokine-cytokine receptor interaction & 31 & 0\\
04620 & Toll-like receptor signaling pathway & 39 & 0\\
04621 & NOD-like receptor signaling pathway & 45 & 0\\
04623 & Cytosolic DNA-sensing pathway & 45 & 0\\
04630 & Jak-STAT signaling pathway & 28 & 0\\
\addlinespace
04210 & Apoptosis & 35 & 0\\
04380 & Osteoclast differentiation & 31 & 0\\
04622 & RIG-I-like receptor signaling pathway & 38 & 0\\
05160 & Hepatitis C & 28 & 0\\
04940 & Type I diabetes mellitus & 35 & 0\\
\bottomrule
\end{tabular}

>>>>

The manuscript was last compile on 2020-08-11 06:50:22.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{BiocManager}\OperatorTok{::}\KeywordTok{install}\NormalTok{(}
    \KeywordTok{c}\NormalTok{(}\StringTok{"NMF"}\NormalTok{, }\StringTok{"ggfortify"}\NormalTok{, }\StringTok{"topGO"}\NormalTok{, }\StringTok{"biomaRt"}\NormalTok{,}
      \StringTok{"KEGGprofile"}\NormalTok{, }\StringTok{"BiocWorkflowTools"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{itemize}
\tightlist
\item
  \texttt{scores}: the matrix of shape \texttt{n\_cluster} x \texttt{n\_genes},
  containing for each gene and each cluster, the goodness of fit as
  described above.
\item
  \texttt{labels}: the labels for all of the genes with a
  sufficiently good goodness-of-fit score.
\item
  \texttt{score\_cutoff}: the cutoff used on \texttt{scores} to determine whether to
  assign a label
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get the best score and best label for all of the genes}
\CommentTok{# This time without filtering labels}
\CommentTok{# We can give the previous calculated scores to `previous_scores` to save time}
\NormalTok{unfiltered_scores =}\StringTok{ }\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(}
\NormalTok{    moanin_model, kmeans_clusters,}
    \DataTypeTok{proportion_genes_to_label=}\DecValTok{1}\NormalTok{,}
    \DataTypeTok{previous_scores=}\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{scores)}
\NormalTok{best_score =}\StringTok{ }\KeywordTok{rowMin}\NormalTok{(scores_and_labels}\OperatorTok{$}\NormalTok{scores)}
\NormalTok{best_label =}\StringTok{ }\NormalTok{unfiltered_scores}\OperatorTok{$}\NormalTok{labels}

@@

\caption{Heatmap of genes in Cluster 6}\label{fig:heatmapCluster8}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  \textbf{Elbow method}. First introduced in 1953 by Thorndike \citep{thorndike:who},
  the elbow methods looks at the total within cluster sum of squares
  as a function of the number of clusters (WCSS). When adding clusters
  doesn't decrease the WCSS by a sufficient
  amount, one can consider stopping. This method thus provides visual aid to
  the user to choose the number of clusters, but often the ``elbow'' is hard to
  see on real data, where the number of clusters is not clearly defined.
\item
  \textbf{Silhouette method}. Similarly to the Elbow method, the Silhouette method
  refers to a method of validation of consistency within clusters, and
  provides visual aid to choose the number of clusters.
\item
  \textbf{Stability methods} Stability methods are more computationally intensive
  than any other method, as they rely on assessing the stability of the
  clustering for every \(k\) to a small randomization of the data. The user is
  then invited to choose the number of cluster based on a number of similarity
  measures.
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{subsample_proportion =}\StringTok{ }\FloatTok{0.8}
\NormalTok{indices =}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\KeywordTok{dim}\NormalTok{(de_moanin_model)[}\DecValTok{1}\NormalTok{],}
\NormalTok{         n_genes }\OperatorTok{*}\StringTok{ }\NormalTok{subsample_proportion,}
         \DataTypeTok{replace=}\OtherTok{FALSE}\NormalTok{)}

@@

\begin{table}[H]
\centering\begingroup\fontsize{7}{9}\selectfont

\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
  & B1 & B2 & B3 & B4 & B5 & B6 & B7 & B8 & B9 & B10 & B11 & B12 & B13 & B14 & B15 & B16 & B17 & B18 & B19 & B20 & B21 & B22 & B23 & B24 & B25\\
\hline
AA204140 & 5 & 5 & 5 & 5 & 5 & 1 & 3 & 2 & 1 & 3 & 1 & 3 & 3 & 1 & 5 & 1 & 1 & 4 & 3 & 5 & 4 & 2 & 4 & 1 & 4\\
\hline
AB008911 & 5 & 5 & 1 & 4 & 1 & 2 & 3 & 3 & 2 & 3 & 3 & 3 & 5 & 1 & 5 & 1 & 3 & 4 & 4 & 5 & 5 & 2 & 4 & 1 & 4\\
\hline
AB010313 & 3 & 3 & 1 & 4 & 1 & 2 & 4 & 3 & 2 & 5 & 3 & 1 & 5 & 3 & 3 & 2 & 3 & 5 & 4 & 2 & 5 & 3 & 3 & 4 & 2\\
\hline
AB010342 & 5 & 3 & 1 & 5 & 1 & 1 & 3 & 3 & 1 & 3 & 3 & 3 & 5 & 1 & 5 & 1 & 1 & 4 & 3 & 5 & 4 & 2 & 3 & 4 & 4\\
\hline
AB011473 & 2 & 1 & 2 & 1 & 3 & 3 & 2 & 5 & 5 & 2 & 4 & 4 & 2 & 5 & 2 & 5 & 5 & 1 & 1 & 1 & 2 & 5 & 1 & 2 & 1\\
\hline
AB099518 & 2 & 1 & 2 & 1 & 3 & 3 & 2 & 5 & 5 & 2 & 4 & 4 & 2 & 5 & 2 & 5 & 5 & 1 & 1 & 1 & 2 & 5 & 1 & 2 & 1\\
\hline
\end{tabular}
\endgroup{}
\end{table}

@@

\begin{table}[H]
\centering\begingroup\fontsize{7}{9}\selectfont

\begin{tabular}{llrr}
\toprule
  & Pathway & Percentage & Adj. p-value\\
\midrule
04060 & Cytokine-cytokine receptor interaction & 31 & 0\\
04620 & Toll-like receptor signaling pathway & 39 & 0\\
04621 & NOD-like receptor signaling pathway & 45 & 0\\
04623 & Cytosolic DNA-sensing pathway & 45 & 0\\
04630 & Jak-STAT signaling pathway & 28 & 0\\
\addlinespace
04210 & Apoptosis & 35 & 0\\
04380 & Osteoclast differentiation & 31 & 0\\
04622 & RIG-I-like receptor signaling pathway & 38 & 0\\
05160 & Hepatitis C & 28 & 0\\
04940 & Type I diabetes mellitus & 35 & 0\\
\bottomrule
\end{tabular}
\endgroup{}
\end{table}
